WI FastPlants: Heritability of number of cotelydons

### 2022-03-19 ms/response2reviews.md

We thank the reviewers for such helpful comments that have greatly improved the web apps and the manuscript.

# Review 1

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- We have added more comments to the code and we have added a new section on the documentation for young audiences interested in coding
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- For the `webinar-aug20` sample data, we make `plant ID` a group variable. We found that for the other web apps, it did not make sense to change `plant ID` as categorical variable because some plots became not very intepretable, so we decided to keep them as quantitiative variables.

add more details on how to install thatssorandom
add more details on how to run runApp
overall check all the steps to run a web app locally
more comments on code! Think of younger audiences looking at the code. No need to comment every line, but explain the overall structure and what each function is doing
“Group variable” should only display the variables that are categorical and “Quantity” should only display the variables that are quantitative
For cotelydon toy data, change plant ID to categorical variable
change the warning on guide scales=FALSE to “none”
check the warnings on fewer than two data points

Add DOI to references
Change Data summary to Data upload
Expand documentation to help students
Make slides available and data for the slides: Maybe change to md
remove the notebooks
editorial comments on broken links

### 2022-03-19 ms/response2reviews.md

We thank the reviewers for such helpful comments that have greatly improved the web apps and the manuscript.

# Review 1

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- We have added more comments to the code and we have added a new section on the documentation for young audiences interested in coding
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- For the `webinar-aug20` sample data, we make `plant ID` a group variable. We found that for the other web apps, it did not make sense to change `plant ID` as categorical variable because some plots became not very intepretable, so we decided to keep them as quantitiative variables.

Software paper
- We added the DOI to me Williams and Hill (1986) paper
- We now make reference to the "Data Upload" section in the paper instead of the "Data Summary"

# Review 2

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- Unfortunately, the `Warning: guides(<scale> = FALSE) is deprecated. Please use guides(<scale> = "none") instead.` is out of our control as we are not calling `guides` directly in our mosaic function. We wondered whether we should suppress all warnings so that the user is not confused, but decided against this in case there are bugs or package upgrades that we need to address in the future. For now, we simply added a section about warnings in the documentation.
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- We have also expanded the documentation including more details about the sample dataset and some of the most common errors and warnings
- The slides provided in the repo are the slides used when teaching the WI Fast plants webinars. The plots in the slides were not created using the same sample data in the repo (but data related to the specific webinar). The goal of the slides was to illustrate concepts, rather than testing the web app. This is actually related to another point brought up by the reviewer (the notebooks). It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, this was not true). In short, the slides are made available as part of the webinars to illustrate statistical concepts rather than usage of the web apps. I have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots. Hope this satisfies the reviewer!

Notebooks folder
- We have double checked the steps to the notebook to make sure they are indeed reproducible and we have included a readme file explaining how the notebooks can be used to reproduce plots and analyses

Software paper
- We have added the DOI to all papers

Editorial
- We fully revised the documentation
- Thanks for checking that the links to the webinars do not work anymore. The webinars were run in coordination with WI Fast Plants and Carolina Knowledge Center, so the links are on their end. We have reached out to them to ask whether the links have changed, but we have not heard back yet.

Expand documentation to help students
  - section on younger audiences starting to code: point at shiny app tutorial
  - section on warnings in the code: Warning: Groups with fewer than two data points have been dropped
  - section on sample data
  - section on most common errors
make the notebooks reproducible (change to sample data)
  - distinguish which are for webinar and which are for reproducing
  - add readme to slides that the slides are meant to illustrate stat concepts

### 2022-03-19 ms/response2reviews.md

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- We have added more comments to the code and we have added a new section on the documentation for young audiences interested in coding
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- For the `webinar-aug20` sample data, we make `plant ID` a group variable. We found that for the other web apps, it did not make sense to change `plant ID` as categorical variable because some plots became not very intepretable, so we decided to keep them as quantitiative variables.

Software paper
- We added the DOI to me Williams and Hill (1986) paper
- We now make reference to the "Data Upload" section in the paper instead of the "Data Summary"

# Review 2

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- Unfortunately, the `Warning: guides(<scale> = FALSE) is deprecated. Please use guides(<scale> = "none") instead.` is out of our control as we are not calling `guides` directly in our mosaic function. We wondered whether we should suppress all warnings so that the user is not confused, but decided against this in case there are bugs or package upgrades that we need to address in the future. For now, we simply added a section about warnings in the documentation.
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- We have also expanded the documentation including more details about the sample dataset and some of the most common errors and warnings
- The slides provided in the repo are the slides used when teaching the WI Fast plants webinars. The plots in the slides were not created using the same sample data in the repo (but data related to the specific webinar). The goal of the slides was to illustrate statistical concepts, rather than testing the web app. This is actually related to another point brought up by the reviewer (the notebooks). It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, they were not working properly). We have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots. We have also added a readme file explaining this in the repo. Hope this satisfies the reviewer!

@@

Notebooks folder
- We have double checked the steps to the notebook to make sure they are indeed reproducible and we have included a readme file explaining how the notebooks can be used to reproduce plots and analyses

Software paper
- We have added the DOI to all papers

Editorial
- We fully revised the documentation
- Thanks for checking that the links to the webinars do not work anymore. The webinars were run in coordination with WI Fast Plants and Carolina Knowledge Center, so the links are on their end. We have reached out to them to ask whether the links have changed, but we have not heard back yet. If it is ok with the reviewers, we will create a new issue on the broken links for now to address this when we receive the new links.

### 2022-03-19 notebooks/cotyledon-data-analysis.Rmd
changes to notebooks

# Initial data: `Polycot-data-from-NY.csv`

@@

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("data/Polycot-data-from-NY.csv", sep=",",header=TRUE)
str(dat)
```

@@

# New data: `Polycot-data-2020.csv`

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
str(dat)
head(dat)
summary(dat)
dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
```

The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
=======
The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.

# New data: `Polycot-data-2020.csv` focusing on chi-square test

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
str(dat)
head(dat)
summary(dat)
dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
```

```{r}
dat2 = data.frame(generation=dat$Generation)
dat2$cots = "dicot"
dat2$cots[dat$Cotyledons > 2] = "polycot"
dat2 = within(dat2, cots<-factor(cots))
summary(dat2)
write.table(dat2,file="data/polycot-data-2020-categories.csv", row.names = FALSE, sep=",")

>>>>

# Data: `Polycot-data-2020.csv` focusing on chi-square test (webinar)

@@

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("../data/Polycot-data-2020.csv", sep=",",header=TRUE)
str(dat)
head(dat)
summary(dat)
dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
```

```{r}
dat2 = data.frame(generation=dat$Generation)
dat2$cots = "dicot"
dat2$cots[dat$Cotyledons > 2] = "polycot"
dat2 = within(dat2, cots<-factor(cots))
summary(dat2)
#write.table(dat2,file="data/polycot-data-2020-categories.csv", row.names = FALSE, sep=",")

# Data: `Polycot-data-2020.csv` (topics not covered in the webinar)

The code of this section contains plots and statistical tests that were not included in the final version of the webinar. For the webinar, we decided to focus on the chi-square test. Here, we fit a t test, a Levene test (for equality of variances) and a Wilcoxon-Mann-Whitney test which is similar to the t test when normality is violated. We also fit an ANOVA test and its alternative when normality is not met, the Kruskal-Wallis rank sum test.

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("../data/Polycot-data-2020.csv", sep=",",header=TRUE)
str(dat)
head(dat)
summary(dat)
dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
```

The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.

# Initial data `Polycot-data-from-NY.csv` (not publicly available)

The following sections have code that was written before the final topics of the webinar were decided. The code below is only kept for internal book-keeping as the data is not made publicly available (this is not the data used in the webinar). Unlike the main section above, here we do not fit a chi-square test, but a t test, a Levene test (for equality of variances) and a Wilcoxon-Mann-Whitney test which is similar to the t test when normality is violated. We also fit an ANOVA test and its alternative when normality is not met, the Kruskal-Wallis rank sum test.

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("../data/Polycot-data-from-NY.csv", sep=",",header=TRUE)
str(dat)
```

@@
